### 2019-07-09 bp_data.tex
Addressed most of the concerns of reviewer 2

Figure~\ref{surveyBio} illustrates the number of survey responses with associated biometric data for each study participant. Participant S2 and S12 have particularly low numbers of usable survey responses. In each of these cases, the issue related to biometric data not being uploaded successfully to the server.

@@

\subsubsection{Response Transformations}
Table~\ref{responseDistribution} illustrates the distribution of responses from each participant for each of the three survey questions (which are listed in Section~\ref{sec:Surveys}). The figure shows that there is a notable imbalance in the distribution of the self-reported responses provided by the participants. Most participants did not use all five points of the five-point Likert scale in their responses, and the distributions tend to skew toward one side or the other, depending on the question. Thus, we binarized the survey data into a two-point scale to give the machine learning models the best possible chance to make useful predictions. The two points in the binary scale represent negative or positive responses for each of the three human aspects of interests (e.g., not stressed or stressed).

@@

\begin{table}
  \centering
      \includegraphics[width=0.5\textwidth]{distributiontable.pdf}
  \caption{The distribution of the responses of each participant to the three questions asked during the day are shown. Each bar in the histograms represent one of the 5 values on the 5-point Likert scale we asked participants to respond with, where the far left side of the histograms are 1/Not at all, and the far right sides are 5/Extremely}
   \label{responseDistribution}
   \vspace*{-9mm}
\end{table}

>>>>

Figure~\ref{surveyBio} illustrates the number of survey responses with associated biometric data for each study participant. 
%This was added to address REviewer 2's concern: Is it correct that the maximum number of responses should have been 112 on the surveys?
The total number of responses per participant is affected by their response rate and by the number days out-of-office (e.g., vacations, holidays, etc.). 
Participant S2 and S12 have particularly low numbers of usable survey responses. In each of these cases, the issue related to biometric data not being uploaded successfully to the server.

@@

\subsubsection{Response Transformations}
Table~\ref{responseDistribution} illustrates the distribution of responses from each participant for each of the three survey questions (listed in Section~\ref{sec:Surveys}). The figure shows that there is a notable imbalance in the distribution of the self-reported responses provided by the participants. Most participants did not use all five points of the five-point Likert scale in their responses, and the distributions tend to skew toward one side or the other, depending on the question. Thus, we binarized the survey data into a two-point scale to give the machine learning models the best possible chance to make useful predictions. The two points in the binary scale represent negative or positive responses for each of the three human aspects of interests (e.g., not stressed or stressed).

@@

\begin{table}
  \centering
      \includegraphics[width=0.5\textwidth]{distributiontable.pdf}
  \caption{The distribution of the responses of each participant to the three questions asked during the day are shown. Each bar in the histograms represent one of the 5 values on the 5-point Likert scale we asked participants to respond with, where the far left side of the histograms are 1/Not at all, and the far right sides are 5/Extremely}
   \label{responseDistribution}
   \vspace*{-2mm}
\end{table}

### 2019-07-09 bp_intro_v3.tex
Addressed most of the concerns of reviewer 2

Fortunately, with recent advances in sensing technologies there is increasing hope that we can accurately and automatically measure such human aspects. A number of studies have already discussed and investigated the use of biometrics to measure aspects such as a person's cognitive and emotional states~\cite{beatty82,Kramer90,Rowe98,richter1998psychophysiological,Wilson2002}. Studies have also looked more specifically at using biometrics to measure stress~\cite{dishman2000stress,Setz2010}, or awakeness / (energetic) arousal~\cite{exler2016tired,mcduff2012affectaura,haag2004emotion}. Yet, very few studies measure productivity-related factors in the workplace over an extended period of time, and little is known about measuring the more abstract concept of focus. Given the importance of focus, described as a combination of engagement and challenge in a work task by Mark et al.~\cite{mark2014bored}, to an office workers' productivity, this factor also deserves further study.

>>>>

Fortunately, with recent advances in sensing technologies there is increasing hope that we can accurately and automatically measure such human aspects. A number of studies have already discussed and investigated the use of biometrics to measure aspects such as a person's cognitive state 
%This change is regarding Reviewer 2's concern: what are the "aspects"?
(e.g., processing load)~\cite{beatty82,Kramer90,richter1998psychophysiological,Rowe98,Wilson2002} 
and more specifically to measure stress~\cite{dishman2000stress,Setz2010}, 
or awakeness / (energetic) arousal~\cite{exler2016tired,mcduff2012affectaura,haag2004emotion}. Yet, very few 
studies measure productivity-related factors in the workplace over an 
extended period of time, and little is known about measuring the more 
abstract concept of focus (directed attention). 
% Added regarding concern of Reviewer 2: Define what is meant by focus when it's first introduced. 
Given the importance of focus, described as a 
combination of engagement and challenge in a work task by Mark et al.~\cite{mark2014bored}, to an office workers' productivity, this factor also 
deserves further study.

### 2019-07-09 bp_relatedWork.tex
addressed concerns from Reviewer 1 using information from their review and our rebuttal
Addressed most of the concerns of reviewer 2

Previous studies have sought to measure, detect, and predict productivity indicators.

>>>>

Previous approaches~\cite{zuger18,Lalle16,Panwar18} have shown the feasibility of applying machine learning 
to biometric data to measure variables such as stress, interruptibility, 
negative emotion, or confusion during computer 
interaction. These approaches vary in the 
variables examined, the invasiveness of the sensors, and the length and 
context of the studies. Many previous studies focused on 
shorter time frames (mostly hours, but in one case also up to five 
days~\cite{sano2013stress} 
or two weeks~\cite{zuger18}). To the best of the authors' knowledge, no 
other study has examined such an extensive period (eight weeks) in a real-
world office environment and the inspected variables (stress, awakeness, and focus).

### 2019-07-09 bp_results_1.tex
Addressed most of the concerns of reviewer 2

\subsection{Predicting Stress, Focus \& Awakeness}
% random forest best
\label{secOverallAccuracy}
As a first step, we compared multiple classifiers using the popular machine learning library scikit-learn~\cite{pedregosa11} and performing a grid search analysis to determine the optimal hyperparameters for each classifier. Our analysis showed that random forest outperforms all other classifiers, including Na\"ive Bayes, decision trees, support vector machine, random forest, and a multilayer perceptron neural network. The optimal values for random forest and the three output measures are listed in Table \ref{tab:hyperparams}. For the remainder of this paper, we will be referring to random forest classifiers trained with these hyperparameters.%\\[-0.1cm]
% for stress they were (minimum samples for split = 4, # of estimators = 100, max features = 0.5, k=300), for focus they were (minimum samples for split = 4, # of estimators = 50, max features = 0.5, k=200), and for awakeness they were (minimum samples for split = 4, # of estimators = 100, max features = 0.25, k=800)
\begin{table}[h]
	\begin{centering}
    \begin{tabular}{llll}
      \hline
      Variable & K & \# Estimators & Minimum Samples Split \\
      \hline
      Stress & 300 & 100 & 4\\
      Focus & 200 & 50 & 4\\
      Awakeness & 800 & 100 & 4\\
      \hline
    \end{tabular}
    \caption{The hyperparameters selected by grid search analysis to tune our random forest models. K refers to the number of features selected.}    \label{tab:hyperparams}
    \end{centering}
\end{table}

@@

Overall, we were able to use the extracted physiological features to predict all three aspects with reasonable accuracy, precision, and recall, as well as to improve upon the baseline---a stratified random classifier that randomly chooses one of the two classes with a bias towards the larger class. While the individually trained classifiers improved on average across all participants upon the baseline in all cases excepting recall of 'not focused', the improvement was substantially higher for awakeness (85\% improvement in precision,  62\% in recall, and 7\% in accuracy) than for stress or focus. Also, the performance of the individually trained classifiers varied greatly across participants. While some participants showed a large improvement, for others the baseline performed much better than the individually trained classifier. For instance, for predicting 'stressed', the individual classifiers improved upon the baseline for S4, S6, S8, S11, S12, and S14 with a maximum improvement of 88.4\% in precision and 184.6\% in recall for S12, while they did worse for S1, S3, S5, S7, S9, S10, and S13, and in the worst cases did not correctly predict a single instance of 'stressed'.\\[-0.1cm]
% We attribute these discrepancies to the large differences in response distributions between participants, as well as to the subjectivity of self-reporting.
%
%(improvement: 8\% precision, 1\% recall, 8\% accuracy), the individual performance varied greatly among participants. Some participants showed negligible difference in comparison to the baseline classifier, e.g. subject ..., while others showed a large improvement, e.g. subject ...
%\begin{table}[h!]
%\begin{centering}
%\begin{tabular}{lll}
%\hline
%Participant & Recall Improvement (\%) & Precision Improvement (\%)\\
%\hline
%S12 & 184.6 & 88.4 \\
%S6 & 66.7 & 85.8 \\
%S11 & 50.0 & 44.4 \\
%S8 & 32.4 & 26.7 \\
%S14 & 14.7 & 9.8 \\
%S4 & 5.6 & 5.5 \\
%S9 & -55.4 & -46.8 \\
%S3 & -90.0 & -82.1 \\
%S1 & -100.0 & -100.0 \\
%S5 & -100.0 & -100.0 \\
%S7 & -100.0 & -100.0 \\
%S10 & -100.0 & -100.0 \\
%S13 & -100.0 & -100.0\\
%S2 & - & -\\
%\hline
%\end{tabular}
%\caption{Percentage improvement in recall and precision for stress, using our approach compared to the baseline, on an individual level. For participant 2, the baseline achieved a precision and recall of 0, thus the improvement is undefined}
%\end{centering}
%\label{tab:indImprovement}
%\end{table}

@@

\begin{table}[h!]
  \begin{centering}
  \begin{tabular}{llll}
    \hline
    Feature Category & Stress & Focus & Awakeness\\
    \hline
    Heart Rate Variability & 18.3\% & 13\% & 13.6\%\\
    Blood Pulse Wave & 10\% & 14.2\% & 13.1\%\\
    Heart Rate & 8.7\% & 12.6\% & 10.3\%\\
    Skin Temperature & 15.7\% & 9.8\% & 10\%\\
    Galvanic Skin Response & 6.6\% & 8.2\% & 5.1\%\\
    Respiration Rate & 14.8\% & 12.7\% & 10\%\\
    Oxygen Saturation & 5.6\% & 4.3\% & 2\%\\ 
    Energy Expenditure & 6\% & 7.7\% & 4.8\%\\
    Activity & 4.6\% & 7.8\% & 7.6\%\\
    Steps & 1.7\% & 0.8\% & 0.9\%\\
    Time of Day & 0\% & 0.1\% & 0.5\%\\
    \hline
  \end{tabular}
  \caption{The averaged Gini importance of each feature category, per response variable.}
  \label{tab:featureImportance}
  \end{centering}
  \vspace*{-5mm}
\end{table}

>>>>

\subsection{Predicting Stress, Focus \& Awakeness}
% random forest best
\label{secOverallAccuracy}
As a first step, we compared multiple classifiers using the popular machine learning library scikit-learn~\cite{pedregosa11} and performing a grid search analysis to determine the optimal hyperparameters for each classifier. Our analysis showed that random forest outperforms all other classifiers, including Na\"ive Bayes, decision trees, support vector machine, random forest, and a multilayer perceptron neural network. The optimal values for random forest and the three output measures are listed in Table \ref{tab:hyperparams}. For the remainder of this paper, we will be referring to random forest classifiers trained with these hyperparameters.%\\[-0.1cm]
% for stress they were (minimum samples for split = 4, # of estimators = 100, max features = 0.5, k=300), for focus they were (minimum samples for split = 4, # of estimators = 50, max features = 0.5, k=200), and for awakeness they were (minimum samples for split = 4, # of estimators = 100, max features = 0.25, k=800)
\begin{table}[h]
	\begin{centering}
	\small\addtolength{\tabcolsep}{-1pt}
    \begin{tabular}{llll}
      \hline
      Variable & K & \# Estimators & Minimum Samples Split \\
      \hline
      Stress & 300 & 100 & 4\\
      Focus & 200 & 50 & 4\\
      Awakeness & 800 & 100 & 4\\
      \hline
    \end{tabular}
    \caption{The hyperparameters selected by grid search analysis to tune our random forest models. K refers to the number of features selected.}    \label{tab:hyperparams}
    \end{centering}
\end{table}

@@

Overall, we were able to use the extracted physiological features to predict all three aspects with reasonable accuracy, precision, and recall, as well as to improve upon the baseline---a stratified random classifier that randomly chooses one of the two classes with a bias towards the larger class. While the individually trained classifiers improved on average across all participants upon the baseline in all cases excepting recall of 'not focused', the improvement was substantially higher for awakeness (85\% improvement in precision,  62\% in recall, and 7\% in accuracy) than for stress or focus. Also, the performance of the individually trained classifiers varied greatly across participants. While some participants showed a large improvement, for others the baseline performed much better than the individually trained classifier. For instance, for predicting 'stressed', the individual classifiers improved upon the baseline for S4, S6, S8, S11, S12, and S14 with a maximum improvement of 88.4\% in precision and 184.6\% in recall for S12, while they did worse for S1, S3, S5, S7, S9, S10, and S13, and in the worst cases did not correctly predict a single instance of 'stressed'. Typically users that have the lowest precision and recall values are those where the data is the most unbalanced. This translates to a scenario where the classifier will likely label the minority instances as part of the majority class due to the unbalanced ratio of the data.\\[-0.1cm]
% We attribute these discrepancies to the large differences in response distributions between participants, as well as to the subjectivity of self-reporting.
%
%(improvement: 8\% precision, 1\% recall, 8\% accuracy), the individual performance varied greatly among participants. Some participants showed negligible difference in comparison to the baseline classifier, e.g. subject ..., while others showed a large improvement, e.g. subject ...
%\begin{table}[h!]
%\begin{centering}
%\begin{tabular}{lll}
%\hline
%Participant & Recall Improvement (\%) & Precision Improvement (\%)\\
%\hline
%S12 & 184.6 & 88.4 \\
%S6 & 66.7 & 85.8 \\
%S11 & 50.0 & 44.4 \\
%S8 & 32.4 & 26.7 \\
%S14 & 14.7 & 9.8 \\
%S4 & 5.6 & 5.5 \\
%S9 & -55.4 & -46.8 \\
%S3 & -90.0 & -82.1 \\
%S1 & -100.0 & -100.0 \\
%S5 & -100.0 & -100.0 \\
%S7 & -100.0 & -100.0 \\
%S10 & -100.0 & -100.0 \\
%S13 & -100.0 & -100.0\\
%S2 & - & -\\
%\hline
%\end{tabular}
%\caption{Percentage improvement in recall and precision for stress, using our approach compared to the baseline, on an individual level. For participant 2, the baseline achieved a precision and recall of 0, thus the improvement is undefined}
%\end{centering}
%\label{tab:indImprovement}
%\end{table}

@@

\begin{table}[h!]
  \begin{centering}
  \begin{tabular}{llll}
    \hline
    Feature Category & Stress & Focus & Awakeness\\
    \hline
    Heart Rate Variability & 18.3\% & 13\% & 13.6\%\\
    Blood Pulse Wave & 10\% & 14.2\% & 13.1\%\\
    Heart Rate & 8.7\% & 12.6\% & 10.3\%\\
    Skin Temperature & 15.7\% & 9.8\% & 10\%\\
    Galvanic Skin Response & 6.6\% & 8.2\% & 5.1\%\\
    Respiration Rate & 14.8\% & 12.7\% & 10\%\\
    Oxygen Saturation & 5.6\% & 4.3\% & 2\%\\ 
    Energy Expenditure & 6\% & 7.7\% & 4.8\%\\
    Activity & 4.6\% & 7.8\% & 7.6\%\\
    Steps & 1.7\% & 0.8\% & 0.9\%\\
    Time of Day & 0\% & 0.1\% & 0.5\%\\
    \hline
  \end{tabular}
  \caption{The averaged Gini importance of each feature category, per response variable.}
  \label{tab:featureImportance}
  \end{centering}
  \vspace*{-2mm}
\end{table}

### 2019-07-09 bp_results_2.tex
Addressed most of the concerns of reviewer 2

\subsection{Minimum Number of Training Samples}\label{secLearningCurve}
Collecting experience samples from users is expensive, since participants are being interrupted several times a day and have to answer the questions. To minimize the number of samples to be collected from participants, we examined how the performance of individual classifiers changes over the number of samples used to train the classifier.

Participants in our study had varying levels of responsiveness to the experience sampling ranging from 10 to 76 (see Figure~\ref{responseDistribution}). To maintain a certain generalizability while also being able to examine a range of sample numbers for training the classifier, we decided to include the ten participants with the most samples for this analysis. As a result, we had at least 34 survey responses per participant to examine the learning curve of the classifiers. For our analysis, we thus performed a leave-one-out cross validation with random sample sets of size 1 to 33 and calculated the average through all folds of the validation.

@@

Figure~\ref{timeWindowsPandR} shows how the F-score changes for predicting `awake' over the 16 different time windows. The figure shows an increasing trend in the F-score, i.e. the higher the number of included time windows included, the higher the F-score. However, there is one exception, the time window of 1200 seconds that achieves a performance close to the one for the time window 10,800 seconds (3 hours), at which point all features are included. Overall, our results thus show that while using all time windows up to 3 hours performs best, and outperforms the feature set that is solely based on a 10 second time window by 28\% (from 0.18 to 0.24), the performance for a time window of 1200 seconds is already very close to optimal (everything up to 3 hours in our case).

@@

\begin{table}
\begin{center}
\begin{tabular}{l l}
\hline

Feature collected by tool & Description \\ 
\hline
Total keystrokes per min& Sum of all types of keystrokes \\ 
Normal keystrokes per min&F[h] Not backspace and navigation \\ 
Backspace keystrokes per min& Backspace keystrokes \\ 
Navigation keystrokes per min& Arrow key keystrokes \\ 
Total clicks per min& Sum of all click types \\ 
Other clicks per min& Not right and left clicks \\ 
Left clicks per min& Left clicks \\ 
Right clicks per min& Right clicks \\ 
Scrolled distance per min& Scrolled distance in pixels \\ 
Moved distance per min& Mouse movements in pixels \\ 
Activity switches per min& Browser window title changes \\ 
Category switches per min& Activity performed category \\ 
\hline
\end{tabular}
\caption{Features collected per user by the computer interaction tracker}%~\cite{meyer18}}
\label{tracker}
\end{center}
\vspace*{-4mm}
\end{table}

@@

Table~\ref{ciPerformance} lists the results of our analysis. The results show that in all cases, the computer interaction based model was able to improve upon the biometric model in terms of precision and recall, but not in accuracy. Further, we found that the combined model was the most effective model in terms of precision and recall for predicting stress and awakeness overall, but performed slightly worse than the model using only computer interaction features for focus. 
%Since we consider precision and recall for predicting the class of higher interest, i.e. stressed, not focused, not awake, to be the most important statistics when interpreting our results, we compared the models against each other by averaging these two statistics.

\begin{table}
\begin{center}
\begin{tabular}{llllll}
\hline
Model/Feature Set & Accuracy & Precision & Recall & F-Score\\
\hline
\textbf{Awakeness}\\
\hspace{3mm}Biometrics only & 0.808 & 0.269 & 0.314 & 0.289\\
\hspace{3mm}C.I. only & 0.758 & 0.425 & 0.362 & 0.391\\
\hspace{3mm}Biometrics + C.I. & 0.791 & 0.390 & 0.404 & 0.400\\
\hline
\textbf{Stress}\\
\hspace{3mm}Biometrics only & 0.775 & 0.270 &	0.260 & 0.265\\
\hspace{3mm}C.I. only & 0.698 & 0.290 & 0.272 & 0.281\\
\hspace{3mm}Biometrics + C.I. & 0.712 & 0.317 & 0.286 & 0.301\\
\hline
\textbf{Focus}\\
\hspace{3mm}Biometrics only & 0.716 & 0.251 & 0.256 & 0.253\\
\hspace{3mm}C.I. only & 0.742 & 0.332 & 0.342 & 0.337\\
\hspace{3mm}Biometrics + C.I. & 0.745 & 0.340 & 0.316 & 0.328\\
\hline
\end{tabular}
\caption{Comparison of the performance of predicting stress, focus and awakeness using the 3 different feature sets for the 10 participants. The performance is calculated as the average of the performance of the individual classifiers. Computer Interactions is abbreviated as C.I. here for readability. Precision and recall refer to the prediction of the more important classes, i.e. `stressed', `not awake', `not focused'.}
\label{ciPerformance}
\end{center}
\vspace*{-6mm}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{lc}
\hline
Model/Feature Set & Number of Features Selected\\
\hline
\textbf{Stress}\\
\hspace{3mm}C.I. & 400\\
\hspace{3mm}Biometrics + C.I. & 800\\
\hline
\textbf{Focus}\\
\hspace{3mm}C.I. & 20\\
\hspace{3mm}Biometrics + C.I. & 300\\
\hline
\textbf{Awakeness}\\
\hspace{3mm}C.I. & All\\
\hspace{3mm}Biometrics + C.I. & 50\\
\hline

\end{tabular}
\caption{The optimal number of features we found to select for each of the model/feature set combinations. Computer interactions is abbreviated as C.I.}
\label{ciFeatureSelection}
\end{center}
\vspace*{-7mm}
\end{table}

>>>>

\subsection{Minimum Number of Training Samples}\label{secLearningCurve}
Collecting experience samples from users is expensive, since participants 
are being interrupted several times a day and have to answer the survey. 
To minimize the number of samples to be collected from participants, we 
examined how the performance of individual classifiers changes over the 
number of samples used to train the classifier.

Participants in our study had varying levels of responsiveness to the 
experience sampling ranging from 10 to 76 (see Figure~
\ref{responseDistribution}).
%Answering Reviewer 2's concern: Why only 10 participants? What was the rationale for the cut-off selected? 
To answer this research question we analyze the maximum number of responses \textbf{all} analyzed participants have to train the model. Since some participants have a very small number of maximum answers (ten being the lowest) we excluded the four participants with the least number of responses for this analysis.
%To maintain a certain generalizability while 
%also being able to examine a range of sample numbers for training the 
%classifier, we removed the four participants with the least number of 
%samples for this analysis. 
As a result, we obtained a corpus of analysis where all participants have a 
higher number of responses (34 being the lowest), which allows us to examine 
the learning curve of the classifiers for a larger number of training data points. For our 
analysis, we thus performed a leave-one-out cross validation with random 
sample sets of size 1 to 33 and calculated the average through all folds of 
the validation.

@@

Figure~\ref{timeWindowsPandR} shows how the F-score changes for predicting 
`awake' over the 16 different time windows. The figure shows an increasing 
trend in the F-score, i.e. the higher the number of included time windows, 
the higher the F-score. However, there is one exception, the time window of 
1200 seconds that achieves a performance close to the one for the time 
window 10,800 seconds (3 hours), at which point all features are included. 
Overall, our results thus show that while using all time windows up to 3 
hours performs best, and outperforms the feature set that is solely based on 
a 10 second time window by 28\% (from 0.18 to 0.24), the performance for a 
time window of 1200 seconds is a good trade-off for selecting a shorter time 
window while maintaining high performance.

@@

\begin{table}
\begin{center}
\small\addtolength{\tabcolsep}{-1pt}
\begin{tabular}{l l}
\hline

Feature collected by tool & Description \\ 
\hline
Total keystrokes per min& Sum of all types of keystrokes \\ 
Normal keystrokes per min&F[h] Not backspace and navigation \\ 
Backspace keystrokes per min& Backspace keystrokes \\ 
Navigation keystrokes per min& Arrow key keystrokes \\ 
Total clicks per min& Sum of all click types \\ 
Other clicks per min& Not right and left clicks \\ 
Left clicks per min& Left clicks \\ 
Right clicks per min& Right clicks \\ 
Scrolled distance per min& Scrolled distance in pixels \\ 
Moved distance per min& Mouse movements in pixels \\ 
Activity switches per min& Browser window title changes \\ 
Category switches per min& Activity performed category \\ 
\hline
\end{tabular}
\caption{Features collected per user by the computer interaction tracker}%~\cite{meyer18}}
\label{tracker}
\end{center}
\vspace*{-1mm}
\end{table}

@@

Table~\ref{ciPerformance} lists the results of our analysis. The results show that in all cases, the computer interaction based model was able to improve upon the biometric model in terms of precision and recall.
%MS is removing the accuracy results because it is not what we care about and it is just making the results harder to understand
%, but not in accuracy. 
Further, we found that the combined model was the most effective model in terms of precision and recall for predicting stress and awakeness overall, but performed slightly worse than the model using only computer interaction features for focus. 
%Since we consider precision and recall for predicting the class of higher interest, i.e. stressed, not focused, not awake, to be the most important statistics when interpreting our results, we compared the models against each other by averaging these two statistics.

\begin{table}
\begin{center}
%\small\addtolength{\tabcolsep}{-1pt}
\begin{tabular}{llllll}
\hline
Model/Feature Set & Precision & Recall & F-Score \\ %& Accuracy\\
\hline
\textbf{Awakeness}\\
\hspace{3mm}Biometrics only  & 0.269 & 0.314 & 0.289 \\ %& 0.808\\
\hspace{3mm}C.I. only  & 0.425 & 0.362 & 0.391 \\ % & 0.758\\
\hspace{3mm}Biometrics + C.I. & 0.390 & 0.404 & 0.400 \\ % & 0.791 \\
\hline
\textbf{Stress}\\
\hspace{3mm}Biometrics only  & 0.270 &	0.260 & 0.265 \\ % & 0.775\\
\hspace{3mm}C.I. only & 0.290 & 0.272 & 0.281 \\ % & 0.698 \\
\hspace{3mm}Biometrics + C.I. & 0.317 & 0.286 & 0.301 \\ % & 0.712\\
\hline
\textbf{Focus}\\
\hspace{3mm}Biometrics only & 0.251 & 0.256 & 0.253 \\ % & 0.716\\
\hspace{3mm}C.I. only & 0.332 & 0.342 & 0.337 \\ % & 0.742\\
\hspace{3mm}Biometrics + C.I. & 0.340 & 0.316 & 0.328 \\ % & 0.745\\
\hline
\end{tabular}
\caption{Comparison of the performance of predicting stress, focus and awakeness using the 3 different feature sets for the 10 participants. The performance is calculated as the average of the performance of the individual classifiers. Computer Interactions is abbreviated as C.I. here for readability. Precision and recall refer to the prediction of the more important classes, i.e. `stressed', `not awake', `not focused'.}
\label{ciPerformance}
\end{center}
\vspace*{-3mm}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{lc}
\hline
Model/Feature Set & Number of Features Selected\\
\hline
\textbf{Stress}\\
\hspace{3mm}C.I. & 400\\
\hspace{3mm}Biometrics + C.I. & 800\\
\hline
\textbf{Focus}\\
\hspace{3mm}C.I. & 20\\
\hspace{3mm}Biometrics + C.I. & 300\\
\hline
\textbf{Awakeness}\\
\hspace{3mm}C.I. & All\\
\hspace{3mm}Biometrics + C.I. & 50\\
\hline
\end{tabular}
\caption{The optimal number of features we found to select for each of the model/feature set combinations. Computer interactions is abbreviated as C.I.}
\label{ciFeatureSelection}
\end{center}
\vspace*{-4mm}
\end{table}

### 2019-07-09 bp_studyMethod.tex
Addressed most of the concerns of reviewer 2

%!TEX root = bioPrediction_main.tex
\section{Field Study}
We conducted an eight-week field study with 14 participants to investigate the feasibility of predicting stress, focus, and awakeness based on biometric signals. On each workday for eight consecutive weeks, each of the 14 participants wore a biometric sensor while performing their normal work tasks and completed two electronic surveys to report their current self-perceived levels of stress, focus, and awakeness. During the eight week study period, a computer activity tracker was installed and active on 10 of the participants' company-issued laptops (note: a few declined for privacy reasons).

@@

\subsubsection{Biometric Sensors}
Figure~\ref{everion} illustrates Biovotion's Everion, which we used to track the biometric signals of the study participants. The Everion is worn on the upper arm and provides continuous monitoring of certain biometric measurements.\footnote{https://biovotion.zendesk.com/hc/en-us/categories/201633909-Everion-Device} Previous studies~\cite{zuger18,sano2013stress,healey2005detecting,wijsman2011towards,zuger2015interruptibility,goyal2017intelligent} have used similar devices~\cite{Okada11,polar,fitbitCharge} to capture psycho-physiological and biometric measurements.

@@

\begin{table}[h!]
\begin{center}
\begin{tabular}{l l}
\hline
Biometric Measurement & Units of Measure \\ 
\hline
\textbf{Physical Activity} & \cite{fox1999,aldana1996}\\
\hspace{3mm}Intensity of motion & (No unit)  \\
\hspace{3mm}Energy Expenditure & Calories per second (cal/s)  \\
\hspace{3mm}Step counter & Steps  \\
\hline
\textbf{Heart} & \cite{haapalainen2010psycho,healey2005detecting,mulder1992measurement,haag2004emotion} \\
\hspace{3mm}Heart rate & Beats per Minute (bpm) \\
\hspace{3mm}Blood pulse wave & (No Unit)  \\
\hspace{3mm}Heart rate variability (RMSSD) & Milliseconds (ms) \\
\hspace{3mm}Blood oxygenation  & Percent (\%)  \\
\hspace{3mm}Blood perfusion & (No unit)  \\
\hline
\textbf{Skin} & \cite{healey2005detecting,haag2004emotion} \\
\hspace{3mm}Galvanic skin response &  kOhm  \\
\hspace{3mm}Skin temperature &  Degrees Celsius ($^{\circ}$C)  \\
\hline
\textbf{Respiration} & \cite{mulder1992measurement,healey2005detecting,haag2004emotion,masaoka1997}\\
\hspace{3mm}Respiratory rate & Breaths per Minute (bpm)  \\
\hline
%\textbf{Environmental}\\
%\hspace{3mm}Barometric pressure & Milibar (mbar)  \\
%\hline
\end{tabular}
\caption{Biometric measurements captured by the Everion, organized by category and with references to previous works using similar data}
\label{signals}
\end{center}
\vspace*{-4mm}
\end{table}

>>>>

%!TEX root = bioPrediction_main.tex
\section{Field Study}
We conducted an eight-week field study with 14 participants to investigate the feasibility of predicting stress, focus, and awakeness based on biometric signals. On each workday for eight consecutive weeks, each of the 14 participants wore a biometric sensor while performing their normal work tasks and completed two electronic surveys to report their current self-perceived levels of stress, focus, and awakeness. During the eight week study period, a computer activity tracker was installed and active on 10 of the participants' company-issued laptops (note: four participants from the original sample declined for privacy reasons).

@@

\subsubsection{Biometric Sensors}
Figure~\ref{everion} illustrates Biovotion's Everion, which we used to track the biometric signals of the study participants. The Everion is worn on the upper arm and provides continuous monitoring of certain biometric measurements.\footnote{https://biovotion.zendesk.com/hc/en-us/categories/201633909-Everion-Device} Previous studies~\cite{zuger18,sano2013stress,healey2005detecting,wijsman2011towards,zuger2015interruptibility,goyal2017intelligent} have used similar devices~\cite{Okada11,polar,fitbitCharge} to capture psycho-physiological and biometric measurements for shorter periods of time or capturing a smaller number of measurements.

@@

\begin{table}[h!]
\begin{center}
%\small\addtolength{\tabcolsep}{-2pt}
\begin{tabular}{l l}
\hline
Biometric Measurement & Units of Measure \\ 
\hline
\textbf{Physical Activity} & \cite{fox1999,aldana1996}\\
\hspace{3mm}Intensity of motion & (No unit)  \\
\hspace{3mm}Energy Expenditure & Calories per second (cal/s)  \\
\hspace{3mm}Step counter & Steps  \\
\hline
\textbf{Heart} & \cite{haapalainen2010psycho,healey2005detecting,mulder1992measurement,haag2004emotion} \\
\hspace{3mm}Heart rate & Beats per Minute (bpm) \\
\hspace{3mm}Blood pulse wave & (No Unit)  \\
\hspace{3mm}Heart rate variability (RMSSD) & Milliseconds (ms) \\
\hspace{3mm}Blood oxygenation  & Percent (\%)  \\
\hspace{3mm}Blood perfusion & (No unit)  \\
\hline
\textbf{Skin} & \cite{healey2005detecting,haag2004emotion} \\
\hspace{3mm}Galvanic skin response &  kOhm  \\
\hspace{3mm}Skin temperature &  Degrees Celsius ($^{\circ}$C)  \\
\hline
\textbf{Respiration} & \cite{mulder1992measurement,healey2005detecting,haag2004emotion,masaoka1997}\\
\hspace{3mm}Respiratory rate & Breaths per Minute (bpm)  \\
\hline
%\textbf{Environmental}\\
%\hspace{3mm}Barometric pressure & Milibar (mbar)  \\
%\hline
\end{tabular}
\caption{Biometric measurements captured by the Everion, organized by category and with references to previous works using similar data}
\label{signals}
\end{center}
\vspace*{-4mm}
\end{table}
